Community Embeddings with Bayesian Gaussian Mixture Model and Variational Inference\\ \thanks{The reported study was partially supported by RFBR grant \textnumero 20-07-00958. The paper was prepared within the framework of the HSE University Project Group Competition 2020-2022.}

### 2020-10-14 paper-ComE-BGMM-VI.tex
rename file
cite . format

Graphs, such as social networks, knowledge graphs, content-rating graphs, and communication networks, emerge naturally from various real-world situations. Analyzing these graphs leads to findings and understanding of the underlying structures, coherences, and dependencies. Recently, methods for embedding graph's nodes into lower-dimensional Euclidean spaces, called graph embeddings, have gained traction in multiple areas of data science research.\cite{Goyal_2018}

Community Embeddings, in addition to embedding a graph's nodes through first- and second-order proximity, also preserve higher-order proximity by embedding clusters present in the graph data. The graph and community embedding algorithm ComE aims to preserve first-, second- and higher-order proximity by embedding a graph's nodes and communities.\cite{ComE} ComE requires prior knowledge of the number of communities $K$. In this paper, ComE is extended to utilizing a Bayesian Gaussian mixture model with variational inference for learning community embeddings (ComE BGMM+VI), similar to ComE+ published by \citeauthor{ComE+} in \citeyear{ComE+}.\cite{ComE+} ComE BGMM+VI takes $K$ as the maximum number of communities and drops components through a trade-off hyperparameter.

The recent \citeyear{ComE} graph embeddings algorithm ComE is extended similarily to the \citeyear{ComE+} ComE+ by taking a Bayesian approach. The open-source code for the Bayesian approach to ComE's community embedding is published on GitHub\footnote{at \citeurl{ComE_BGMM_GH}} and serves as a contribution to community embedding research.\cite{ComE_BGMM_GH}

The original ComE paper \citetitle{ComE} by \citeauthor{ComE} and the ComE+ paper \citetitle{ComE+} by \citeauthor{ComE+} in combination with the ComE source code have provided the basis and architecture for the community embeddings utilized in this work.\cite{ComE, ComE+, ComE_GH}

Multiple surveys and articles on graph embeddings were consulted to build a full picture of the current state of graph embedding research. Especially the \citeyear{Goyal_2018} survey \citetitle{Goyal_2018} by \citeauthor{Goyal_2018} and the \citeyear{rossi20tkdd-roles} paper \citetitle{rossi20tkdd-roles} by \citeauthor{rossi20tkdd-roles} have proven to be primary resources for understanding the current landscape of graph embedding research.\cite{Goyal_2018, rossi20tkdd-roles}

On part of comparing the Bayesian Gaussian mixture model with variational inference to the Gaussian mixture model with expectation-maximization, the \citeyear{Bishop06} book \citetitle{Bishop06} by \citeauthor{Bishop06} includes essential statistics and information science knowledge and explanations.\cite{Bishop06}

>>>>

Graphs, such as social networks, knowledge graphs, content-rating graphs, and communication networks, emerge naturally from various real-world situations. Analyzing these graphs leads to findings and understanding of the underlying structures, coherences, and dependencies. Recently, methods for embedding graph's nodes into lower-dimensional Euclidean spaces, called graph embeddings, have gained traction in multiple areas of data science research \cite{Goyal_2018}.

Community Embeddings, in addition to embedding a graph's nodes through first- and second-order proximity, also preserve higher-order proximity by embedding clusters present in the graph data. The graph and community embedding algorithm ComE aims to preserve first-, second- and higher-order proximity by embedding a graph's nodes and communities\cite{ComE}. ComE requires prior knowledge of the number of communities $K$. In this paper, ComE is extended to utilizing a Bayesian Gaussian mixture model with variational inference for learning community embeddings (ComE BGMM+VI), similar to ComE+ published by \citeauthor{ComE+} in \citeyear{ComE+} \cite{ComE+}. ComE BGMM+VI takes $K$ as the maximum number of communities and drops components through a trade-off hyperparameter.

The recent \citeyear{ComE} graph embeddings algorithm ComE is extended similarily to the \citeyear{ComE+} ComE+ by taking a Bayesian approach. The open-source code for the Bayesian approach to ComE's community embedding is published on GitHub\footnote{at \url{https://github.com/abegehr/ComE_BGMM}} and serves as a contribution to community embedding research \cite{ComE_BGMM_GH}.

The original ComE paper \textit{Learning Community Embedding with Community Detection and Node Embedding on Graphs} by \citeauthor{ComE} and the ComE+ paper \textit{Embedding Both Finite and Infinite Communities on Graphs} by \citeauthor{ComE+} in combination with the ComE source code have provided the basis and architecture for the community embeddings utilized in this work \cite{ComE, ComE+, ComE_GH}.

Multiple surveys and articles on graph embeddings were consulted to build a full picture of the current state of graph embedding research. Especially the \citeyear{Goyal_2018} survey \textit{Graph Embedding Techniques, Applications, and Performance: A Survey} by \citeauthor{Goyal_2018} and the \citeyear{rossi20tkdd-roles} paper \textit{On Proximity and Structural Role-based Embeddings in Networks: Misconceptions, Techniques, and Applications} by \citeauthor{rossi20tkdd-roles} have proven to be primary resources for understanding the current landscape of graph embedding research \cite{Goyal_2018, rossi20tkdd-roles}.

On part of comparing the Bayesian Gaussian mixture model with variational inference to the Gaussian mixture model with expectation-maximization, the \citeyear{Bishop06} book \textit{Pattern Recognition and Machine Learning (Information Science and Statistics)} by \citeauthor{Bishop06} includes essential statistics and information science knowledge and explanations \cite{Bishop06}.

\section{Community Embedding}
